Fractional Types

### 2016-12-19 typespp.tex
fig:apply
fig:compose
trace

\unitetl/ :& \forall\exists{\overline{k}}.[\order{\idiso} \circledast \mathbb{T},

@@

The adaptation of the combinators to the pointed dependent setting in Fig .~\ref
{pifracd:comb} is relatively straightforward. First, the types of the
combinators must keep track of how the value currently in focus changes during
evaluation, and hence directly encode the operational semantics of the
language. As suggested above, the quantifiers are interpreted differently
depending on which direction the rule is used. We use a $\forall\exists$
notation to express the presence of \emph{one} quantifier that is interpreted
ddifferently in each direction. The notation has the advantage of being compact,
avoiding duplication, and making the symmetry of the rules explicit. We also
include two new pairs of combinators $\synchl/packl$ and $\synchr/packr$ that
serve as explicit quantifier introduction and elimination rules: they serve as
explicit ``synchronization'' points as illustrated in the example below.

@@

\paragraph*{Forward execution of zigzag circuit with synchronization.}
\[\begin{array}{rcl}
[\order{c},c^i]
&\stackrel{\unitetl/}{\leftrightblackspoon}&
  [\order{id}\circledast\order{c}, \langle \idiso,c^i \rangle] \\
&\stackrel{\eta+\circledast\idiso/}{\leftrightblackspoon}&
  \forall{k}.[(\order{c}\circledast\iorder{c})\circledast\order{c},
  \langle \langle c^k,1/c^k \rangle,c^i \rangle] \\
&\stackrel{\synchl}{\leftrightblackspoon}&
  \langle \langle c^i,1/c^i \rangle,c^i \rangle] \\
&\stackrel{\assocrt/}{\leftrightblackspoon}&
  [\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^i,c^i \rangle\rangle] \\
&\stackrel{\packr}{\leftrightblackspoon}&
  \exists{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^k,c^k \rangle\rangle] \\
&\stackrel{\idiso/\circledast\epsilon-}{\leftrightblackspoon}&
  [\order{c}\circledast\order{\idiso},
  \langle c^i, \idiso \rangle] \\
&\stackrel{\unitetr/}{\leftrightblackspoon}&
  [\order{c}, c^i]
\end{array}\]

@@

>>>>

\unitetl/ :& \mbox{\AE}{\overline{k}}.[\order{\idiso} \circledast \mathbb{T},

@@

The adaptation of the combinators to the pointed dependent setting in Fig .~\ref
{pifracd:comb} is relatively straightforward. First, the types of the
combinators must keep track of how the value currently in focus changes during
evaluation, and hence directly encode the operational semantics of the
language. As suggested above, the quantifiers are interpreted differently
depending on which direction the rule is used. We use a $\forall\exists$
notation to express the presence of \emph{one} quantifier that is interpreted
differently in each direction. The notation has the advantage of being compact,
avoiding duplication, and making the symmetry of the rules explicit. We also
include two new pairs of combinators $\synchl/packl$ and $\synchr/packr$ that
serve as explicit quantifier introduction and elimination rules: they serve as
explicit ``synchronization'' points as illustrated in the example below.

@@

\paragraph*{Forward execution of zigzag circuit with synchronization.}
\[\begin{array}{rcl}
[\order{c},c^i]
&\stackrel{\unitetl/}{\leftrightblackspoon}&
  [\order{id}\circledast\order{c}, \langle \idiso,c^i \rangle] \\
&\stackrel{\eta+\circledast\idiso/}{\leftrightblackspoon}&
  \forall{k}.[(\order{c}\circledast\iorder{c})\circledast\order{c},
  \langle \langle c^k,1/c^k \rangle,c^i \rangle] \\
&\stackrel{\synchl}{\leftrightblackspoon}&
  [\langle \langle c^i,1/c^i \rangle,c^i \rangle] \\
&\stackrel{\assocrt/}{\leftrightblackspoon}&
  [\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^i,c^i \rangle\rangle] \\
&\stackrel{\packr}{\leftrightblackspoon}&
  \exists{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^k,c^k \rangle\rangle] \\
&\stackrel{\idiso/\circledast\epsilon-}{\leftrightblackspoon}&
  [\order{c}\circledast\order{\idiso},
  \langle c^i, \idiso \rangle] \\
&\stackrel{\unitetr/}{\leftrightblackspoon}&
  [\order{c}, c^i]
\end{array}\]

@@

%%%
\subsection{Trace}

Traced categories~\cite{joyal1996traced} have proved useful for modeling
recursion~\cite{Hasegawa:1997:RCS:645893.671607}. It is interesting that every
compact closed category admits a trace which we can implement as shown below. In
the figure, we show that given a combinator
$\rho_\bullet : \order{c} \circledast \order{d} \leftrightblackspoon \order{c}
\circledast \order{e}$
we can construct a circuit of type $\order{d} \leftrightblackspoon \order{e}$
that typechecks iff we have a fixed point for $f$ for the argument of type
$\order{c}$.

FIX TRACE

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}[scale=0.9,every node/.style={scale=0.7},highlight/.style={line width=4pt,orange!70}]
    \path [highlight] (-1,1) edge node[above] {} (0,1);
    \path (-1,1) edge node[above] {\footnotesize{$\order{d}$}} (0,1);

    \path [highlight] (9,1.5) edge node[above] {} (10,1.5);
    \path (9,1.5) edge node[above] {\footnotesize{$1/{\textsf{\#c}}$}} (10,1.5);
  \end{tikzpicture}
  \caption{\label{fig:trace} Trace}
\end{figure}
